Donor Preferences in Canadian Political Financing

### 2024-12-04 paper/paper.qmd
update appendix

\newpage
# Data {#sec-data}

@@

\newpage
# Model {#sec-model}

@@

@tbl-modelsummary presents the summary of key model estimates for
the Ontario, Federal, and Conservative/Liberal models. These models
predict the logarithm of total donations received by political parties
based on whether the party was in power, the size of the party, and
whether the donation occurred during an election year. The
Conservative/Liberal model includes an interaction term to explore how
the effect of being in power varies between these two major parties.

@@

## Weaknesses and next steps
This study provides valuable insights into donation behaviours across Canada but it its limitations. First, the analysis is applied to only federal and provincial donation records, overlooking regional variations found in other provinces and territories. The geographical limitations of this study hinder its ability to generalize across the entire country. Moreover, the study only spans from 2006 to 2024 which economic or political events could have influence over the donation patterns that the models do not account for. Additionally qualitative factors such as personal beliefs, specific policy preferences and campaign effectiveness are not directly addressed in this study. Data quality issues such as potential missing values or inconsistencies could also have affected the accuracy of the findings here. Lastly, the linear regression models here assume a linear relationship between predictors and donations which can greatly oversimplify the complexities.

@@

The dataset used in this analysis underwent a rigorous data cleaning
process to ensure accuracy, consistency, and reliability. The raw
donation records were sourced from the Investigative Journalism
Foundation (IJF), encompassing political donations at both the federal
and Ontario provincial levels from 2013 to 2024. The data cleaning
process involved several critical steps, detailed below.

Initially, the dataset was examined for missing or incomplete values in
key variables such as Amount Donated, Political Party, Donation Year,
and Region. Records with missing values in these essential fields were
excluded from the analysis to maintain data integrity and prevent biases
that could arise from incomplete data entries.

Political party names often appeared in various formats due to
inconsistencies in reporting. To address this, party names were
standardized across all records. For instance, variations such as
"Liberal Party of Ontario" and "Ontario Liberal Party" were consolidated
under the standardized name "Ontario Liberal Party." This
standardization ensured accurate attribution of donations to the correct
parties.

Donor information, including names and locations, was also standardized
where possible. Typographical errors were corrected, and formatting was
made consistent (e.g., "First Name Last Name" instead of "Last Name,
First Name"). However, original donor names were retained in cases where
standardization could introduce inaccuracies, particularly when dealing
with individuals with similar names.

The dataset included both monetary (amount_monetary) and non-monetary
(amount_non_monetary) contributions. A new variable, Amount Donated, was
created by summing these two fields for each record. This provided a
comprehensive measure of the total value of donations received by each
party. All monetary values were converted to a consistent currency
format, and any discrepancies in currency symbols or formats were
corrected to ensure accurate aggregation and comparison of donation
amounts.

To facilitate analysis at the party-year level, individual donation
records were aggregated. Donations were summed for each political party
within each year, resulting in a dataset where each entry represented
the total donations received by a party in a specific year. This
aggregation allowed for the examination of trends over time and
assessment of the impact of factors such as power status and election
years on donation amounts.

@@

Recipient_in_Power: Indicates whether the recipient party was in power
during the donation year (1 for in power, 0 for not in power).
Party_Size: Categorizes parties as major (1) or minor (0) based on
legislative representation and historical influence. Election_Year:
Indicates whether the donation was made during an election year (1 for
election year, 0 for non-election year). These variables were crucial
for the regression analysis, allowing for the assessment of the
influence of these factors on total donations.

The distribution of total donation amounts was highly skewed, with a
small number of large donations and many smaller ones. To address this
skewness and stabilize variance, a logarithmic transformation was
applied to the Total Donations variable, creating Log_Total_Donations.
This transformation also helped in handling zero values by adding a
small constant (e.g., 1) before taking the logarithm.

To ensure the accuracy of the cleaned dataset, cross-validation was
performed against original records and official financial statements
where available. This step involved spot-checking aggregated totals and
individual records to confirm that the cleaning and aggregation
processes did not introduce errors.

Some challenges were encountered due to inconsistencies in the raw data,
such as typographical errors, inconsistent reporting formats, and
missing donor types. While efforts were made to standardize and correct
these issues, some limitations remain due to the quality of the original
data sources. In cases where donor type information was missing and
legal restrictions indicated that only individuals could donate (e.g.,
Ontario after 2017), the donor type was inferred as "Individual." While
this assumption is reasonable based on the legal context, it may not
capture all nuances in the data.

@@

The Investigative Journalism Foundation (IJF) employs a comprehensive
approach to collect and compile political donation data across Canada.
Understanding the IJF's data collection methodology is essential for
evaluating the reliability and scope of the data used in this analysis.

**Data Sources and Acquisition**

@@

-   Parties: Registered political parties at the federal and provincial
    levels.

@@

Contribution limits and the types of entities permitted to donate vary
across jurisdictions. For instance, at the federal level, only
individuals who are Canadian citizens or permanent residents can donate,
with maximum annual contributions set at specified amounts. In contrast,
some provinces allow donations from corporations, unions, and other
organizations and may have different or no contribution limits.

@@

Extensive manual cleaning is performed to correct errors arising from
OCR conversion and to standardize the data. Key steps in the data
cleaning process include:

@@

Original records submitted by political entities may contain
typographical errors or inaccuracies, which are reflected in the
dataset. Incomplete or Inconsistent Data: Variations in reporting
requirements and practices across jurisdictions can lead to
inconsistencies in the data. OCR Limitations: Despite manual
verification, OCR technology may not perfectly capture handwritten or
poorly scanned documents, potentially introducing errors. Conclusion

### Idealized Methodology

@@

A stratified random sampling technique will be utilized to ensure
proportional representation from various demographics and regions. The
sample will be stratified based on:

@@

To maintain a national margin of error of ±3% at a 95% confidence level,
a total sample size of approximately 1,500 respondents is required. This
sample size allows for meaningful analysis of subgroups within the data.
Given an estimated response rate of 10%, outreach efforts will target
approximately 15,000 potential respondents across Canada.

@@

To incentivize participation, all respondents will be offered a small
monetary compensation (e.g., a \$10 electronic gift card) or entered
into a national sweepstakes with a chance to win one of ten \$500 cash
prizes.

@@

-   nclude attention-check questions to identify inattentive or
    fraudulent responses.

@@

    -   This covers costs associated with designing the survey,
        including question formulation, structuring for clarity and
        flow, pilot testing, and ensuring compliance with ethical
        standards. Additionally, funds account for bilingual translation
        services (English and French).

@@

Timeline

@@

### Idealized Survey Design

Survey Structure

[Purpose of survey]
You can access the survey here: https://docs.google.com/forms/d/e/1FAIpQLSdj2RxS9fjZ_n6Q54zjSyoJ7tAJT7ssuTF3V1tVk6a_Rhyp4w/viewform?usp=sharing

@@

**State of Residence:** - Dropdown list of all provinces and
territories.

**Age:** Multiple-choice options

@@

**Gender:**

@@

**Income Level:**

@@

**Education Level:** Highest degree obtained

@@

**Political Affiliation:**

@@

**Are you currently registered to donate to political parties?**

@@

**How frequently do you donate to political parties?**

@@

**Average Donation Amount:**

@@

**Preferred Donation Channels:**

@@

**What motivates you to donate to a political party?** (Select all that
apply)

@@

**Which factors most influence your decision to donate?** (Select all
that apply)

@@

**Have recent political events influenced your donation behavior?**

@@

**If yes, please specify the events and how they influenced your
donations.** (Open-ended)

**How do you perceive the effectiveness of political campaigns in
encouraging donations?**

@@

**What could political parties do to encourage more donations from
supporters?** (Open-ended) -

**Any additional comments or suggestions regarding political
donations?** (Open-ended)

>>>>

# Data {#sec-data}

@@

# Model {#sec-model}

@@

@tbl-modelsummary presents the summary of key model estimates for the
Ontario, Federal, and Conservative/Liberal models. These models predict
the logarithm of total donations received by political parties based on
whether the party was in power (In_Power), the size of the party
(Party_Size), and whether the donation occurred during an election year
(Election_Year). The Conservative/Liberal model includes an interaction
term (In_Power × Party) to explore how the effect of being in power
varies between these two major parties.

@@

## Weaknesses and next steps

This study provides valuable insights into donation behaviours across
Canada but it its limitations. First, the analysis is applied to only
federal and provincial donation records, overlooking regional variations
found in other provinces and territories. The geographical limitations
of this study hinder its ability to generalize across the entire
country. Moreover, the study only spans from 2006 to 2024 which economic
or political events could have influence over the donation patterns that
the models do not account for. Additionally qualitative factors such as
personal beliefs, specific policy preferences and campaign effectiveness
are not directly addressed in this study. Data quality issues such as
potential missing values or inconsistencies could also have affected the
accuracy of the findings here. Lastly, the linear regression models here
assume a linear relationship between predictors and donations which can
greatly oversimplify the complexities.

@@

Rigorous data cleaning was done to the dataset used in this
investigation to assure correctness and dependability. The raw donation
records were sourced from the Investigative Journalism Foundation (IJF)
which include political donations at both federal and provincial levels,
specifically Ontario, from 2013 to 2024. The details of the data
cleaning processed are described below.

First, missing and incomplete values were searched for in key variables
such as Amount Donated, Donation Year, Political Party and Region.
Excluding any records with missing values in these crucial fields to
minimise biases created from incomplete data entries and to maintain
data integrity.

Inconsistencies in reporting caused political party names to appear in
different formats. To fix this, party names were given constant and
standardized names across all the records. For example, if there was a
variation such as "Liberal Party of Ontario" and "Ontario Liberal Party"
these were combined into a single name "Ontario Liberal Party" to ensure
accurate attribution of donations to parties.

Fields that were also standardized were the donor information which
includes names and locations. Any typographical errors were corrected,
and formatting was made consistent, for example the structure "First
Name Last Name" instead of "Last Name, First Name". But, in cases where
standardization could introduce inaccuracies, original donor names were
kept to avoid dealing with individuals with similar names.

Two columns, (amount_monetary) and (amount_non_monetary) which
correspond to monetary and non-monetary contributes are included in the
data set. Using these a new variable was created that sums these two
fields together for every record. In doing this, the dataset now
provides a comprehensive measure of the total value of donations
received by a party. Moreover, monetary values were converted into a
consistent currency format to disregard any discrepancies in the
currency symbols or formatting and to ensure accurate aggregation and
proper comparison.

Individual donation records were aggregated in order to perform proper
analysis at the party-year level. For each political party within each
year donations were summed yielding a dataset where each entry
represented the total donations received by a party for a specific year.
Aggregating the data like this enabled the examination of trends over
time and impact assessment factors like power status and election years
on donation amounts.

@@

Recipient_in_Power: Indicates whether the recipient party was in power
during the donation year (1 for in power, 0 for not in power).
Party_Size: Categorizes parties as major (1) or minor (0) based on
legislative representation and historical influence. Election_Year:
Indicates whether the donation was made during an election year (1 for
election year, 0 for non-election year). These variables were imperative
for the regression analysis and allowed the assessment of the influence
of these on total donations.

The distribution of total donation amounts was highly skewed, with a
small number of large donations and many smaller ones. In order to
address this, a logarithmic transformation was applied to the
Total_Donations variable creating a new variable called
Log_Total_Donations. This transformation also helped with handling zero
values by adding a small constant before taking the logarithm.

Cross-validation was used against original records and official
financial statements to ensure accuracy of the cleaned dataset. This
step required spot-checking aggregated totals and individuals records to
confirm that the cleaning and aggregation process did not create more
errors.

There were challenges that occurred due to inconsistencies in the raw
data, these included typographical errors, reporting format errors and
missing donor types. Some limitations remained despite the effort to
standardize and correct these issues, due to the quality of the original
data source. When donor type information was missing and legal
restrictions indicated that only individuals could donate, the donor
type was inferred as "Individual". Despite this assumption being legally
reasonable in the context of the paper, it may not capture all nuances
in the data.

@@

The Investigative Journalism Foundation (IJF) employs a comprehensive
approach to collect and compile political donation data across Canada.
Understanding the IJF's data collection methodology is crucial for
assessing the reliability and scope of the data used in this analysis.

Data Sources and Acquisition

@@

-   Parties: Federal and provincial registered political parties.

@@

Jurisdictions have varying limits on the donation amount and types of
entities allowed to donate. For example, in the federal level, only
individuals who are Canadian citizens or permanent residents can donate,
with a maximum annual contribution amount. Comparatively, there are some
provinces that allow donations from corporations, organizations and
unions and may have different or no contribution limits.

@@

Manual cleaning is extensively done to correct errors that arise from
OCR conversion and to standardize the data. The key steps are:

@@

Records submitted by political entities can contain typographical errors
or inaccuracies which are seen in the dataset. Inconsistent or
incomplete data can occur in reporting practices across different
jurisdictions. Moreover, OCR technology is imperfect at capturing
handwritings or poorly scanned documents introducing errors despite
manual correction.

**Idealized Methodology**

@@

A stratified random sampling technique will be utilized to ensure
proportional representation from various demographics and regions. The
sample will be stratified according to:

@@

To maintain a national margin of error of plus or minus 3% at a 95%
confidence level, a total sample size of approximately 1,500 respondents
is required. This sample size allows for meaningful analysis of
subgroups within the data. Given an estimated response rate of 10%,
outreach efforts will target approximately 15,000 potential respondents
across Canada.

@@

To incentivize participation, all respondents will be entered into a
national sweepstakes with a chance to win one of ten \$500 cash prizes.

@@

-   Include attention-check questions to identify inattentive or
    fraudulent responses.

@@

    -   This covers costs associated with designing the survey,
        including question formulation, structuring for clarity and
        flow, pilot testing, and ensuring compliance with ethical
        standards. Additionally, funds account for bilingual translation
        services (English and French). This also covers the cost of the
        potential monetary gifts given after completing the survey.

@@

**Timeline**

@@

#### Idealized Survey Design

**Survey Structure**

@@

State of Residence:

-   Dropdown list of all provinces and territories.

Age: Multiple-choice options

@@

Gender:

@@

Income Level:

@@

Education Level: Highest degree obtained

@@

Political Affiliation:

@@

Are you currently registered to donate to political parties?

@@

How frequently do you donate to political parties?

@@

Average Donation Amount:

@@

Preferred Donation Channels:

@@

What motivates you to donate to a political party? (Select all that
apply)

@@

Which factors most influence your decision to donate? (Select all that
apply)

@@

Have recent political events influenced your donation behavior?

@@

If yes, please specify the events and how they influenced your
donations. (Open-ended)

How do you perceive the effectiveness of political campaigns in
encouraging donations?

@@

What could political parties do to encourage more donations from
supporters? (Open-ended) -

Any additional comments or suggestions regarding political donations?
(Open-ended)
